Example gene set analysis: bringing it together

### 2021-12-07 examples/example1.Rmd
tidy1

@@

dat

barplot(dat,ylab="jaccard metric")

saveRDS(dat,file = "ex1dat.rds")

>>>>

s1 %>% kbl(caption = "sample sheet") %>% kable_paper("hover", full_width = F)

@@

dat
saveRDS(dat,file = "ex1dat.rds")

par(mar=c(5,10,3,1))
barplot(rev(dat),xlab="Jaccard index",horiz = TRUE, las =1, xlim=c(0,.75) , main=name)
text( x=rev(dat)+0.05 , y= 1:length(rev(dat))*1.2-0.5, labels = signif(rev(dat),2))

pdf("images/jacbar.pdf",width=4,height=3)
par(mar=c(5,9,3,1))
barplot(rev(dat),xlab="Jaccard index",horiz = TRUE, las =1, xlim=c(0,.8) , main=name)
text( x=rev(dat)+0.08 , y= 1:length(rev(dat))*1.2-0.5, labels = signif(rev(dat),2))
dev.off()

png("images/jacbar.png")
par(mar=c(5,10,3,1))
barplot(rev(dat),xlab="Jaccard index",horiz = TRUE, las =1, xlim=c(0,.75) , main=name)
text( x=rev(dat)+0.05 , y= 1:length(rev(dat))*1.2-0.5, labels = signif(rev(dat),2))
dev.off()

### 2021-12-07 examples/example6.Rmd
tidy1

---
title: "Example gene set analysis: The case of Set7 knockdown"
author: "Mark Ziemann & Kaumadi Wijesooriya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 5
theme: cosmo
---

@@

The dataset being used is SRP038101 and we are comparing the cells expressing a set7kd shRNA construct (case) compared to the scrambled construct (control).

@@

I'm using some RNA-seq data looking at the effect of Set7 knockdown on HMEC cells.

@@

name="SRP096177"
mdat<-getDEE2Metadata("hsapiens")
samplesheet <- mdat[grep("SRP096177",mdat$SRP_accession),]
samplesheet<-samplesheet[order(samplesheet$SRR_accession),]
samplesheet$trt<-as.factor(c(1,1,1,0,0,0))
s1 <- samplesheet

@@

w<-getDEE2("hsapiens",samplesheet$SRR_accession,metadata=mdat,legacy = TRUE)
x<-Tx2Gene(w)
x<-x$Tx2Gene

@@

png("images/fcs_ora4.png")
plot(euler(v1),quantities = TRUE, edges = "gray", main="FCS vs ORA")
dev.off()

png("images/orabg4.png")
plot(euler(v2),quantities = TRUE, edges = "gray", main="Effect of inappropriate background* (whole genome)")
dev.off()

png("images/oracomb4.png")
plot(euler(vx),quantities = TRUE, main="combining up and down genes and whole genome bg*")
dev.off()

pdf("images/fcs_ora4.pdf",width=4,height=4)
plot(euler(v1),quantities = TRUE, edges = "gray", main="FCS vs ORA")
dev.off()

pdf("images/orabg4.pdf",width=4,height=4)
plot(euler(v2),quantities = TRUE, edges = "gray", main="Effect of inappropriate background* (whole genome)")
dev.off()

pdf("images/oracomb4.pdf",width=4,height=4)
plot(euler(vx),quantities = TRUE, edges = "gray", main="combining up and down genes and whole genome bg*")
dev.off()

@@

saveRDS(dat,file = "ex4dat.rds")

>>>>

---
title: "Example gene set analysis: The case of SARS-CoV2 infection"
author: "Mark Ziemann & Kaumadi Wijesooriya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 5
theme: cosmo
---

@@

The dataset being used is SRP253951 and we are comparing A549 cells with and without infection with SARS-CoV-2.

@@

I'm using some RNA-seq data looking at the difference in transcriptome caused by SARS-CoV-2 infection.

@@

name="SRP253951"
mdat<-getDEE2Metadata("hsapiens")
samplesheet <- mdat[grep("SRP253951",mdat$SRP_accession),]

samplesheet <- mdat[which(mdat$SRX_accession %in% c("SRX8089264","SRX8089265","SRX8089266","SRX8089267","SRX8089268","SRX8089269")),]

samplesheet$trt <- factor(c(0,0,0,1,1,1))
s1 <- samplesheet

@@

w<-getDEE2("hsapiens",samplesheet$SRR_accession,metadata=mdat,legacy = TRUE)
x<-Tx2Gene(w)
x <- x$Tx2Gene

@@

png("images/fcs_ora6.png")
plot(euler(v1),quantities = TRUE, edges = "gray", main="FCS vs ORA")
dev.off()

png("images/orabg6.png")
plot(euler(v2),quantities = TRUE, edges = "gray", main="Effect of inappropriate background* (whole genome)")
dev.off()

png("images/oracomb6.png")
plot(euler(vx),quantities = TRUE, main="combining up and down genes and whole genome bg*")
dev.off()

pdf("images/fcs_ora6.pdf",width=4,height=4)
plot(euler(v1),quantities = TRUE, edges = "gray", main="FCS vs ORA")
dev.off()

pdf("images/orabg6.pdf",width=4,height=4)
plot(euler(v2),quantities = TRUE, edges = "gray", main="Effect of inappropriate background* (whole genome)")
dev.off()

pdf("images/oracomb6.pdf",width=4,height=4)
plot(euler(vx),quantities = TRUE, edges = "gray", main="combining up and down genes and whole genome bg*")
dev.off()

@@

saveRDS(dat,file = "ex6dat.rds")

### 2021-12-07 examples/example7.Rmd
tidy1

---
title: "Example gene set analysis: The case of SARS-CoV2 infection"
author: "Mark Ziemann & Kaumadi Wijesooriya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 5
theme: cosmo
---

@@

The dataset being used is SRP253951 and we are comparing A549 cells with and without infection with SARS-CoV-2.

@@

I'm using some RNA-seq data looking at the difference in transcriptome caused by SARS-CoV-2 infection.

@@

name="SRP253951"
mdat<-getDEE2Metadata("hsapiens")
samplesheet <- mdat[grep("SRP253951",mdat$SRP_accession),]

samplesheet <- mdat[which(mdat$SRX_accession %in% c("SRX8089264","SRX8089265","SRX8089266","SRX8089267","SRX8089268","SRX8089269")),]

samplesheet$trt <- factor(c(0,0,0,1,1,1))
s1 <- samplesheet

@@

w<-getDEE2("hsapiens",samplesheet$SRR_accession,metadata=mdat,legacy = TRUE)
x<-Tx2Gene(w)
x <- x$Tx2Gene

@@

png("images/fcs_ora6.png")
plot(euler(v1),quantities = TRUE, edges = "gray", main="FCS vs ORA")
dev.off()

png("images/orabg6.png")
plot(euler(v2),quantities = TRUE, edges = "gray", main="Effect of inappropriate background* (whole genome)")
dev.off()

png("images/oracomb6.png")
plot(euler(vx),quantities = TRUE, main="combining up and down genes and whole genome bg*")
dev.off()

pdf("images/fcs_ora6.pdf",width=4,height=4)
plot(euler(v1),quantities = TRUE, edges = "gray", main="FCS vs ORA")
dev.off()

pdf("images/orabg6.pdf",width=4,height=4)
plot(euler(v2),quantities = TRUE, edges = "gray", main="Effect of inappropriate background* (whole genome)")
dev.off()

pdf("images/oracomb6.pdf",width=4,height=4)
plot(euler(vx),quantities = TRUE, edges = "gray", main="combining up and down genes and whole genome bg*")
dev.off()

@@

saveRDS(dat,file = "ex6dat.rds")

>>>>

---
title: "Example gene set analysis: The case of P300 knockdown"
author: "Mark Ziemann & Kaumadi Wijesooriya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 5
theme: cosmo
---

@@

The dataset being used is SRP068733 and we are comparing the healthy endothelial cells with a scrambled siRNA to cells treated with a p300 targeting siRNA.

@@

I'm using some RNA-seq data looking at the difference in healthy endothelial cell gene expression between vehicle and C646 samples.

@@

name="SRP068733"
mdat<-getDEE2Metadata("hsapiens")
samplesheet <- mdat[grep("SRP068733",mdat$SRP_accession),]
samplesheet<-samplesheet[order(samplesheet$SRR_accession),]
SRRvec <- c("SRR3112216","SRR3112217","SRR3112218","SRR3112219","SRR3112220","SRR3112221")
samplesheet <- samplesheet[which(samplesheet$SRR_accession %in% SRRvec),]
samplesheet$trt<-as.factor(c(0,0,0,1,1,1)) 
s1 <- samplesheet

@@

w<-getDEE2("hsapiens",SRRvec,metadata=mdat,legacy = TRUE)
x<-Tx2Gene(w)
x<-x$Tx2Gene

@@

png("images/fcs_ora7.png")
plot(euler(v1),quantities = TRUE, edges = "gray", main="FCS vs ORA")
dev.off()

png("images/orabg7.png")
plot(euler(v2),quantities = TRUE, edges = "gray", main="Effect of inappropriate background* (whole genome)")
dev.off()

png("images/oracomb7.png")
plot(euler(vx),quantities = TRUE, main="combining up and down genes and whole genome bg*")
dev.off()

pdf("images/fcs_ora7.pdf",width=4,height=4)
plot(euler(v1),quantities = TRUE, edges = "gray", main="FCS vs ORA")
dev.off()

pdf("images/orabg7.pdf",width=4,height=4)
plot(euler(v2),quantities = TRUE, edges = "gray", main="Effect of inappropriate background* (whole genome)")
dev.off()

pdf("images/oracomb7.pdf",width=4,height=4)
plot(euler(vx),quantities = TRUE, edges = "gray", main="combining up and down genes and whole genome bg*")
dev.off()

@@

saveRDS(dat,file = "ex7dat.rds")

### 2021-12-07 examples/examples.Rmd
tidy1

means <- colMeans(exdat)
medians <- colMedians(exdat)
sds <- apply(exdat,2,sd)
summstats <- rbind(means,medians,sds)
summstats %>% kbl(caption="summary statistics of different enrichment analysis comparisons") %>% kable_paper("hover", full_width = F)

@@

Now plot as a beeswarm.

```{r,jacswarm}

exdat <- rbind(exdat,"mean"=colMeans(exdat))

cols <- c("#CC79A7","#56B4E9","#E69F00","#F0E442","#009E73","black")
coll <- list(cols, cols, cols, cols, cols)
par(mar=c(5,10,3,1))
beeswarm(list(exdat[,5],exdat[,4],exdat[,3],exdat[,2],exdat[,1]),labels = rev(colnames(exdat)),horizontal = TRUE, las=1 , pwcol = coll,cex=2,pch=19, xlab = "Jaccard index")
legend("bottomright", legend=c("SRP128998", "SRP038101", "SRP096178","SRP038101","SRP247621","mean"), pch=19,  col=cols, cex=1,pt.cex = 1.5)

pdf("images/jacswarm1.pdf",width=6,height=5)
par(mar=c(5,9,3,1))
beeswarm(list(exdat[,5],exdat[,4],exdat[,3],exdat[,2],exdat[,1]),labels = rev(colnames(exdat)),horizontal = TRUE, las=1 , pwcol = coll,cex=2,pch=19, xlab = "Jaccard index")
legend("bottomright", legend=c("SRP128998", "SRP038101", "SRP096178","SRP038101","SRP247621","mean"), pch=19,  col=cols, cex=1.0, pt.cex = 1.7)
dev.off()

png("images/jacswarm1.png")
par(mar=c(5,10,3,1))
beeswarm(list(exdat[,5],exdat[,4],exdat[,3],exdat[,2],exdat[,1]),labels = rev(colnames(exdat)),horizontal = TRUE, las=1 , pwcol = coll,cex=2,pch=19, xlab = "Jaccard index")
legend("bottomright", legend=c("SRP128998", "SRP038101", "SRP096178","SRP038101","SRP247621","mean"), pch=19,  col=cols, cex=1,pt.cex = 1.5)
dev.off()

rbind(exdat,mysd) %>% kbl() %>% kable_paper("hover", full_width = F)

>>>>

means <- colMeans(exdat)
medians <- apply(exdat,2,median)
sds <- apply(exdat,2,sd)
summstats <- rbind(means,medians,sds)
summstats %>% kbl(caption="summary statistics of different enrichment analysis comparisons") %>% kable_paper("hover", full_width = F)

@@
